Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study

### 2020-08-21 paper/aimusic2020/paper.tex
Reduce to 8 pages

\renewcommand{\baselinestretch}{0.96}
\pagestyle{headings}

@@

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{2}\and Helene-Camille Crayencour\inst{3} \thanks{Please place acknowledgement here.}}
%
% if the names of the authors are too long for the running head, please use the format: AuthorA et al.
\authorrunning{Carmine Emanuele Cella et al.}

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{University of California, Berkeley \and InstituteB\\ \email{author@phdsrus.ac.uk}}

@@

\begin{abstract}
In this paper we will do a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration. We will show how it is possible to model this  musical problem as a classification task and we will propose two deep learning models. We will show, first, how they perform as classifiers for musical instrument recognition by comparing them with specific baselines. We will then show how they perform, both qualitatively and quantitatively, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems. Finally, we will highlight benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

@@

Target-based computer-assisted orchestration can be thought of as the process of searching for the best combinations of sounds in a database of instrumental samples, to match a target sound under a specific similarity metric and specific sets of constraints. A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound. 
Valid orchestral scores may contain several instruments sounding simultaneously, selected out of a large number of possible combinations of sounds from the database.

@@

In this paper, we try a different approach to this problem by experimenting with deep neural architectures. The main idea is to train a model to classify combinations of real instruments and use it then for orchestration. A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}. By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation. Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes.

%\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.

This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07, Kitahara05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16, Bian19}.

In this paper we do preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}. We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.

After training these models on varying combinations of 10 instruments, we used them for orchestration. More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system. We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means. For quantitative comparison, we computed a specific distance metric on the amplitude spectrum of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.

Next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results. Sections \ref{sec:orchestration} and \ref{sec:conclusions} will present, respectively, the results of orchestration experiments using the trained models and the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.

@@

A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample. For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.

In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.

Three baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes (the instruments in TinySOL) or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g. C, D\#, G, etc.) for a total of 25 classes (we will classifiy combinations of 2 instruments for 12 pitch classes each, plus one extra \emph{trash} class, see section \ref{sec:baseline}).

Two neural classifiers, instead, have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e. the neural classifiers would classify instruments and pitches with respective octaves, hence 12 instruments for 12 pitch classes and three dynamics; note that not all the samples in TinySOL have all dynamics). We want to point out, however, that this is still a simplified problem compared to assisted orchestration. Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note. To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models. When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}). By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.  Figure \ref{fig:method} illustrates the described ideas.

@@

For a given number of instruments $N$, each input to our models is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments. The data is generated in by selecting $N$ random TinySOL samples, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures. Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments and the Mel frequency cepstral coefficients (MFCCs) and the Mel spectrogram of the mixture are computed to be used as the input features for the baseline classifiers and the neural classifiers respectively. The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.

The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals. The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. %The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
Such setup proved to be successful in \cite{Salamon17}.

@@

The former has been directly applied on the TinySOL samples each time they are selected to generate a new combination. We performed a small pitch shift by reading the samples with different sample rates: a small difference in sample rate will slightly modify the duration and the perceived pitch if played at the normal sample rate. In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.

The latter was performed on the feature matrix itself of input samples, the Mel spectrogram. We chose random columns and rows of the matrix to zero out. For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out. This method of data augmentation aimed to be more resilient to the possible variations in the recording of the instruments.

@@

In order to get a better sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN). We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}. More specifically, for SVM we used a non-linear RBF kernel and for RF we set the maximum depth of each tree to be 15.  All of the baseline experiments used 50,000 generated samples with a train-test split of 60/40. Each sample was made by a combination of one to ten instruments and was four seconds in length. 
In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a lower number of features, that is more manageable by parametric classifiers. We found SVM to have the highest accuracy of the three classifiers across all experiments.

We started with a simplification of our problem. We performed experiments in which two instruments were selected, and the classifier attempted to identify instrument and pitch \emph{class} of the samples. The pitch class is the note name without the octave, e.g. C, D\#, G, etc. The best accuracies from these experiments were from a combination of Oboe and French Horn, in which SVM had an accuracy of 39.9\% and RF had an accuracy of 17.5\%. Since RF performed worse than SVM in every experiment, we stopped testing with it and used SVM from that point on.

We then performed this same experiment with three instruments having their pitch class identified. Flute, Oboe, and Violin reached an accuracy of 11.1\%, and Bass Tuba, Trumpet, and Trombone was 0.5\%. As we increased the number of instruments whose pitch classes was being identified, the accuracy continued to drop.Indeed, for classifying the pitch class of four instruments (Oboe, French Horn, Violin, and Flute) the accuracy was 2.7\%.

This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments. However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased. Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments.

@@

In the rest of this section, we will show the classification results of the two neural models.

@@

The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the one successfully used in \cite{Salamon17}. The LSTM unit was added in order to provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.

@@

Following the first three convolutional layers, there is an LSTM layer which outputs 32 matrices. After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$. We flatten the outputs and feed them into a fully connected layer with Dropout, then another another fully connected layer. Finally, the sigmoid function is applied to the final layer.

Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. %The model architecture is shown in Figure \ref{cnn_lstm}.

@@

The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet}. Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block. Besides, to make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.

@@

Different experiments were made by varying the number $N$ of samples in each mixture. We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb.

Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs. 
%The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was 	trained on 400,000 generated samples over 20 epochs. Fig. \ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
ResNet outperforms the CNN regardless of the number of samples used in the combination. This result is consistent with previous research \cite{He15}, as residual networks usually perform well in classification problems.

@@

After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration. For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet. Each class, in this model, represents an instrument-pitch pair.

@@

Since we decided not to train our models to classify the dynamics of a sample (despite TinySOL having \emph{pianissimo}, \emph{mezzoforte}, and \emph{fortissimo} recordings), the dynamics are determined by the probability of each sample as output by the model. If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample is used. If the probability is between $0.33$ and $0.66$, then the mezzoforte version is selected and if it is less than $0.33$ the pianissimo version is used. The idea behind this quantization is that samples that are the most represented in the target should appear as the loudest in the orchestrated solution.

In order to test our models for orchestration, we used 15 targets from the Orchidea distribution. These targets represent a variety of signal types but are mostly static, in the sense that they do not change sensibly over time.

Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of bassoon and two symphonic chords. The other targets there are: the recordings of two bells, of a car horn, of a gong, of a screaming voice, of a wind harp, of the attack of a brass instrument and the recording of a boat docking.

Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length since we adapt the representation to the size of the training data by changing the hop length during the computation of the Mel spectrogram.

\begin{table*}
  \centering
     \caption{Quantitative comparison of orchestrations as ratios to Orchidea. Eqn. \ref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets. What is shown is the ratio between the distance calculated and Orchidea's distance for the same target. A value less than 1 means that the model performed worse (i.e. had a larger distance), and a value greater than 1 means the model performed better than Orchidea. The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
	\centering    
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
     \hline
      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass cl.} & \textbf{Bell 1} & \textbf{Bell 2} & \textbf{Multiph. 1} & \textbf{Car horn} & \textbf{Boat} $\ldots$ \\
      \hline
      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} $\ldots$  \\
      \hline
      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} $\ldots$ \\
      \hline
      \hline
      \textbf{$\ldots$} & \textbf{Wind harp} & \textbf{Chord 1} & \textbf{Multiph. 2} & \textbf{Chord 2} & \textbf{Gong} & \textbf{Scream} & \textbf{Brass} & \textbf{Average}  \\
      \hline
      $\ldots$ & 0.55 & 0.79 & 0.70 & 0.57 & 0.73 & \textbf{1.14} & 0.79 & 0.71\\
      \hline
      $\ldots$  & 0.61 & 0.86 & 0.51 & 0.37 & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.66 \\
      \hline
    \end{tabular}

@@

We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to the solutions generated by Orchidea, the state-of-the-art system for computer-assisted orchestration.

%Orchidea implements many advanced features that are not supported by our models. For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles. It is also able to reduce the search space by applying harmonic analysis on the target sound, and create sparse solutions that do not use every instrument specified.

In order to have a fairer comparison, we did not allow Orchidea to use any of its advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution. %This creates a fairer comparison, since our models are unable to create constrained or sparse solutions.

Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch. For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution. For example, one of the samples of a bell had partials that loosely represented a C minor chord, so we checked if the orchestration contained the notes of the chord.

For quantitative evaluation, we used the distance metric defined in Eqn. \ref{distance} to calculate differences in timbre between targets and solutions. This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization. The equation takes in the full amplitude spectrum of the target $x$ and full amplitude spectrum of the solution $\hat{x}$. Then for each bin $k$ of the amplitude spectrum, it calculates the absolute difference between the values. The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.

We calculated the distance between the target samples and our orchestrated solutions. We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions. A comparison of these results is in Table \ref{orch_eval}.

@@

While our model is not able to outperform Orchidea, it shows consistent results. We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets. Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target. You can listen to the targets and orchestrated solutions from Orchidea, ResNet, and the CNN with LSTM at \url{https://dzluke.github.io/DeepOrchestration/}.

@@

The first natural extension would be to support sparsity in our models. Our current models, indeed, orchestrate all targets using a constant number of instruments and it is not able to drop specific instruments from the solution. This does not take into account the density of different targets. Sparse solutions, in which the model decides how many samples should be used to best represent the target, would allow a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds. This would generate more meaningful orchestrations that would compare more favourably to the state of the art.

>>>>

\renewcommand{\baselinestretch}{0.96}
\raggedbottom
\pagestyle{headings}

@@

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{2}\and Helene-Camille Crayencour\inst{3}}
%
% if the names of the authors are too long for the running head, please use the format: AuthorA et al.
\authorrunning{Carmine Emanuele Cella et al.}

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{University of California, Berkeley \and InstituteB\\ \email{ , dz.luke@berkeley.edu, }}

@@

\begin{abstract}
In this paper we will perform a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration. We will show how it is possible to model this  musical problem as a classification task and we will propose two deep learning models. We will show, first, how they perform as classifiers for musical instrument recognition by comparing them with specific baselines. We will then show how they perform, both qualitatively and quantitatively, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems. Finally, we will highlight benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

@@

Target-based computer-assisted orchestration can be thought of as the process of searching for the best combinations of sounds in a database of instrumental samples, to match a target sound under a specific similarity metric and specific sets of constraints. A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.

@@

In this paper, we try a different approach to this problem by experimenting with deep neural architectures. The main idea is to train a model to classify combinations of real instruments and then use it for orchestration. A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}. By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation. Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes. This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07, Kitahara05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16, Bian19}.

In this paper we perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}. We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.

@@

A complete orchestration solution, however, would normally be made by triples of \textit{instrument-pitch-dynamics}. It is difficult to frame this problem as classification, since we would need to have a very high number of classes. Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample. For this reasons, our models were not trained to determine \textit{instrument-pitch-dynamics} triples but instead \textit{instrument-pitch} pairs.

@@

For a given number of instruments $N$, each input to our model is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments. The data is generated by selecting $N$ random TinySOL samples, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures. Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments. The Mel spectrogram of the mixture is computed to be used as the input features for the classifiers. The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.

The choice of using the Mel spectrogram as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals. %The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. %The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
Such setup proved to be successful in \cite{Salamon17}.

@@

Pitch shifting was applied on the TinySOL samples each time they were selected to generate a new combination. We performed a small pitch shift by reading the samples with different sample rates: a small difference in sample rate will slightly modify the duration and the perceived pitch if played at the normal sample rate. In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.

Partial feature dropout was performed on the feature matrix itself of input samples, the Mel spectrogram. We chose random columns and rows of the matrix to zero out. %For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out. 
This method of data augmentation aimed to be more resilient to the possible variations in the recording of the instruments.

@@

In order to get a better sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN). We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}. 
In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; as the number of features is more manageable for parametric classifiers. We found SVM to have the highest accuracy of the three classifiers across all experiments.

We started with a simplification of our problem. We performed experiments in which two instruments were selected, and the classifier attempted to identify instrument and pitch \emph{class} of the samples. For two instruments, SVM had an accuracy of 39.9\% and RF had an accuracy of 17.5\%. As the number of instruments used in combination increased, the accuracy dropped sharply. With three instruments, SVM accuracy was 11.1\%, with four instruments it was 2.7\%. Clearly, these classifiers were not going to be able to achieve meaningful results with the full setting of our problem.

@@

%In the rest of this section, we will show the classification results of the two neural models.

@@

The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the success in \cite{Salamon17}. The LSTM unit was added in order to provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.

@@

Following the first three convolutional layers, there is an LSTM layer which outputs 32 matrices. After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$. We flatten the outputs and feed them into a fully connected layer with Dropout, then another another fully connected layer. Finally, the sigmoid function is applied to the final layer. Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. %The model architecture is shown in Figure \ref{cnn_lstm}.

@@

The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet}. Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block. To make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.

@@

Different experiments were made by varying the number $N$ of samples in each mixture. We used an orchestra of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb. Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs. 
%The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was 	trained on 400,000 generated samples over 20 epochs. Fig. \ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
ResNet outperforms the CNN regardless of the number of samples used in the combination. This result is consistent with previous research \cite{He15}, as residual networks usually perform well in classification problems.

@@

After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration.

@@

Since we decided not to train our models to classify the dynamics of a sample, the dynamics are determined by the probability of each sample as output by the model. If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample is used. A probability between $0.33$ and $0.66$, is mezzoforte and less than $0.33$ is pianissimo. The idea behind this quantization is that samples that are the most represented in the target should appear as the loudest in the orchestrated solution.

In order to test our models for orchestration, we used 15 targets from the Orchidea distribution. These targets represent a variety of signal types but are mostly static, in the sense that they do not change sensibly over time. Some of the targets were made of instrumental samples and chords, others are bells or gongs, and some do not feature any musical instruments

\begin{table*}
  \centering
     \caption{Quantitative comparison of orchestrations as ratios to Orchidea. Eqn. \ref{distance} was used to compute distances between orchestrations and targets. What is shown is the ratio between the distance calculated and Orchidea's distance for the same target. A value less than 1 means that the model performed worse (i.e. had a larger distance), and a value greater than 1 means the model performed better than Orchidea. The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
	\centering    
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
     \hline
      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass cl.} & \textbf{Bell 1} & \textbf{Bell 2} & \textbf{Multiph. 1} & \textbf{Car horn} & \textbf{Boat} $\ldots$ \\
      \hline
      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} $\ldots$  \\
      \hline
      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} $\ldots$ \\
      \hline
      \hline
      \textbf{$\ldots$} & \textbf{Wind harp} & \textbf{Chord 1} & \textbf{Multiph. 2} & \textbf{Chord 2} & \textbf{Gong} & \textbf{Scream} & \textbf{Brass} & \textbf{Average}  \\
      \hline
      $\ldots$ & 0.55 & 0.79 & 0.70 & 0.57 & 0.73 & \textbf{1.14} & 0.79 & 0.71\\
      \hline
      $\ldots$  & 0.61 & 0.86 & 0.51 & 0.37 & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.66 \\
      \hline
    \end{tabular}

@@

We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to the solutions generated by Orchidea, the state-of-the-art system for computer-assisted orchestration. 
In order to have a fairer comparison, we did not allow Orchidea to use any of its advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it to use all 10 instruments in each solution.

%Orchidea implements many advanced features that are not supported by our models. For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles. It is also able to reduce the search space by applying harmonic analysis on the target sound, and create sparse solutions that do not use every instrument specified.
%This creates a fairer comparison, since our models are unable to create constrained or sparse solutions.

Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch. For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.

For quantitative evaluation, we used the distance metric defined in Eqn. \ref{distance} to calculate differences in timbre between targets and solutions. This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.

@@

A comparison of distances between our solutions and targets and Orchidea's solutions and targets is in Table \ref{orch_eval}. While our model is not able to outperform Orchidea, it shows consistent results. We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets. Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target. You can listen to the targets and orchestrated solutions from Orchidea, ResNet, and the CNN with LSTM at \url{https://dzluke.github.io/DeepOrchestration/}.

@@

The first natural extension would be to support sparsity in our models. Our current models orchestrate all targets using a constant number of instruments and are not able to drop specific instruments from the solution. This does not take into account the density of different targets. Sparse solutions, in which the model decides how many samples should be used to best represent the target, would allow a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds. %This would generate more meaningful orchestrations that would compare more favourably to the state of the art.
